---
title: "Exploring Weather Trends"
author: "Joao Martins"
output:
  tufte::tufte_html: default
---

```{r setup, include=FALSE}
library(reticulate)
use_condaenv("r-reticulate", required = TRUE)
```

# Outline

To compare annual average temperature records from Portugal against corresponding world annual average temperature records between 1750 and 2013,  we take the following steps:

1. Download regional and global temperature trends from [http://berkeleyearth.lbl.gov/](http://berkeleyearth.lbl.gov) using the command line (see `Bash` script in annex);

2. Load, prepare, and visualize the trends using `Python` (see script in annex):  
  2.1. Open, clean the CSV table with [`pandas`](https://github.com/pandas-dev/pandas);  
  2.2. Produce a line chart showing a moving average of the temperatures with [`plotnine`](https://github.com/has2k1/plotnine)^[`plotnine` is based on `Matplotlib` and implements `ggplot2`'s grammar of graphics, well-suited for exploratory analyses.]. `plotnine`'s [moving average smoothing method](https://plotnine.readthedocs.io/en/stable/generated/plotnine.stats.stat_smooth.html) is based on the `pandas.rolling()` function.

3. Display the results in [R Markdown](https://rmarkdown.rstudio.com), following the [`tufte`](https://github.com/rstudio/tufte) layout.

# Observations

```{r update, include=FALSE, eval=FALSE}
system("../data/get_data.sh", show.output.on.console = TRUE)
# move .csv files to ../data/
```

```{r prep, include=FALSE}
py_run_file("../python/prep.py")
```

```{python mainfigure, echo=FALSE, results='hide', fig.keep='all', fig.align='center', fig.cap='Average temperatures (ºC) between 1750 and 2013.  Each dot represents the average temperature over 1 year. Lines show a moving average temperature over a time period of 10 years.  Temperatures in Portugal are shown in red, world temperatures in blue.'}
p
```

Figure 1 shows a number of interesting observations:

1. Over the past 250 years, Portugal has been consistently warmer than the rest of the world by about 7ºC.

2. Year-over-year temperatures appear to oscillate more at a regional level.  The ruggedness of the moving average and the level of dispersion of the data points seems higher for Portugal.  By contrast, world average temperatures appear to typically more consistent from a year to the next.  Possibly, they do not suffer from regional fluctuations.

3. The 20th century appears to show a regional and global warming trend.


# Annex

```{r sources, include=FALSE}
src_bash <- c(
  "```{bash, eval=FALSE}", 
  readLines("../data/get_data.sh"), 
  "```")
src_python <- c(
  "```{python, eval=FALSE}", 
  readLines("../python/prep.py"), 
  "```")
```

## Bash

`r paste(knitr::knit(text = src_bash), collapse = "\n")`

## Python

`r paste(knitr::knit(text = src_python), collapse = "\n")`
